Issues with Dynamical Measurements Across the Disk

### 2019-02-26 paper/ms.tex
Minor macro changes

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
We briefly describe the cosmological-hydrodynamical zoom-in simulations used
in this work. Cosmological zoom-ins allow one to simulate a selected region at
high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm{}
for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{at \url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
and \mm , respectively. These values are summarized in
Table~\ref{tab:scale_height}.

@@

Cosmological simulations of Milky Way-mass galaxies are not perfect
representations of the true Milky Way in other ways as well, as discussed in
\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi\ is
closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
this work we are most interested in the global properties of the potential,
and specifically in deviations from axisymmetry. From this perspective, the
simulated galaxies are actually \emph{more} axisymmetric than we might expect
of the Milky Way. While they have prominent spiral arms, none has as strong a
bar as the MW does at present day, and none has a nearby companion like the
Large Magellanic Cloud. One of the three we consider (\mf) does have an
ongoing interaction with a satellite galaxy similar to Sagittarius, which has
punched through the galactic disk outside the solar circle and induced some
warping. However, we expect that, as with the use of finite smoothing lengths,
the degree of midplane variation in the simulations is likely a factor of a
few lower than what could be expected in the MW.

@@

To allow for potential small inaccuracies in the determination of the original
fiducial coordinate system, we also subtract the best fit $\text{A}
\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
function of azimuth to account for an overall tilt of the midplane (a
simplified version of the strategy described in
\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-174$, $62$,
and $3.8\,\pc$, for $\text{B}$ the values are $0.3$, $0.65$, and
$0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$, and
$-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for \mi , \mf , and \mm . These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
midplane our imaginary observer would determine as a function of azimuth for
each galaxy, using their local ``solar neighborhood'' (the cylinder defined
above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
shown as the dashed-line error bars. The $90\%$ interquartile range for each
galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the
three cases the midplane therefore varies by more than $\pm 100$pc depending
on the azimuth along the Solar circle; in the third (\mm , which has the
thinnest ``thin disk'' of stars, but the largest stellar mass) the variation
is more like $\pm 50$ pc.

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars within
a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

For the thick-disk orbit (brown), the neighborhood is at the maximum value for
nearly all masses, except at the very low-mass end in \mi .

We repeat this entire procedure for each of the orbits of the real open
clusters in Table~\ref{tab:real_clusters} using the midplane variations from
\mi . We use \mi\ as a conservative choice, since it provides the greatest
midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
values for the mass of each cluster in the neighborhood column of
Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
analysis assumes that the midplane from \mi\ can be accurately applied to
Milky Way orbits, which is not necessarily true. However, our analysis is
meant to be more illustrative than quantitative and these neighborhood numbers
should be understood as such.

\begin{figure*}
\plotone{fig/Rn_vs_mc.pdf}
\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
and thick orbits (\thincolor and \thickcolor, respectively; see
Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
\mi , \mf , and \mm\ ({\em left}, {\em center}, and {\em right} panels,
respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
the ``neighborhood'' around the cluster, i.e. the distance from the cluster
one could go before the $J_z$ error induced by our midplane effect is
comparable to the intrinsic action dispersion of the cluster. When the $z$
offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
procedure but using the $\pm1\sigma$ lines from
Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
is negligible. However, for the thin orbit (\thincolor), which is closest to
the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
neighborhood is only a few hundred $\pc$ for low-mass clusters.}
\label{fig:Rn_mc}
\end{figure*}

@@

\item To demonstrate the previous point, we measured the local Galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (\mi , \mf , and \mm , respectively).

>>>>

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
We briefly describe the cosmological-hydrodynamical zoom-in simulations used
in this work. Cosmological zoom-ins allow one to simulate a selected region at
high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{at \url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
\mf{}, and \mm{}, respectively. These values are summarized in
Table~\ref{tab:scale_height}.

@@

Cosmological simulations of Milky Way-mass galaxies are not perfect
representations of the true Milky Way in other ways as well, as discussed in
\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is
closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
this work we are most interested in the global properties of the potential,
and specifically in deviations from axisymmetry. From this perspective, the
simulated galaxies are actually \emph{more} axisymmetric than we might expect
of the Milky Way. While they have prominent spiral arms, none has as strong a
bar as the MW does at present day, and none has a nearby companion like the
Large Magellanic Cloud. One of the three we consider (\mf) does have an
ongoing interaction with a satellite galaxy similar to Sagittarius, which has
punched through the galactic disk outside the solar circle and induced some
warping. However, we expect that, as with the use of finite smoothing lengths,
the degree of midplane variation in the simulations is likely a factor of a
few lower than what could be expected in the MW.

@@

To allow for potential small inaccuracies in the determination of the original
fiducial coordinate system, we also subtract the best fit $\text{A}
\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
function of azimuth to account for an overall tilt of the midplane (a
simplified version of the strategy described in
\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-174$, $62$,
and $3.8\,\pc$, for $\text{B}$ the values are $0.3$, $0.65$, and
$0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$, and
$-15\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the
values given in \citet{2018arXiv180610564S} for the difference between the
$z$-axis as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
midplane our imaginary observer would determine as a function of azimuth for
each galaxy, using their local ``solar neighborhood'' (the cylinder defined
above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
shown as the dashed-line error bars. The $90\%$ interquartile range for each
galaxy is $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
three cases the midplane therefore varies by more than $\pm 100$pc depending
on the azimuth along the Solar circle; in the third (\mm{}, which has the
thinnest ``thin disk'' of stars, but the largest stellar mass) the variation
is more like $\pm 50$ pc.

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine error bars
(95\uth and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

For the thick-disk orbit (brown), the neighborhood is at the maximum value for
nearly all masses, except at the very low-mass end in \mi{}.

We repeat this entire procedure for each of the orbits of the real open
clusters in Table~\ref{tab:real_clusters} using the midplane variations from
\mi{}. We use \mi{} as a conservative choice, since it provides the greatest
midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
values for the mass of each cluster in the neighborhood column of
Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
analysis assumes that the midplane from \mi{} can be accurately applied to
Milky Way orbits, which is not necessarily true. However, our analysis is
meant to be more illustrative than quantitative and these neighborhood numbers
should be understood as such.

\begin{figure*}
\plotone{fig/Rn_vs_mc.pdf}
\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
and thick orbits (\thincolor and \thickcolor, respectively; see
Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
\mi{}, \mf{}, and \mm{} ({\em left}, {\em center}, and {\em right} panels,
respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
the ``neighborhood'' around the cluster, i.e. the distance from the cluster
one could go before the $J_z$ error induced by our midplane effect is
comparable to the intrinsic action dispersion of the cluster. When the $z$
offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
procedure but using the $\pm1\sigma$ lines from
Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
is negligible. However, for the thin orbit (\thincolor), which is closest to
the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
neighborhood is only a few hundred $\pc$ for low-mass clusters.}
\label{fig:Rn_mc}
\end{figure*}

@@

\item To demonstrate the previous point, we measured the local Galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
